---
title: "Reproducible Research: Peer Assessment Part 1"
author: "Gran Ville Lintao"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.



## Introduction

We set the options and include libraries in this part of code:
```{r setoptions,echo=TRUE, message=FALSE}
library(dplyr)
library(ggplot2)
```



## Loading and preprocessing the data

1. Loading the data can be done via the following code:

```{r}
path2csv <- "./activity.csv"
mydf <- read.csv(path2csv)
```

2. This code converts the data frame to a data frame table structure for easier processing:

```{r}
mytbldf <- tbl_df(mydf)
```



## What is mean total number of steps taken per day?

<!--
```{r}

#mean_by_day <- summarize(by_day, #meansteps=mean(steps))
```
-->

The total number of steps taken per day can be calculated by using the verbs from dplyr - group_by and summarize

```{r}
by_day <- group_by(mytbldf, date)
sum_by_day <- summarize(by_day, sumsteps=sum(steps))

```

And then visualized in a histogram by using ggplot2

code:
```{r}

# convert the date to a vector of integers so it can be displayed nicely in the plot
days <- as.numeric(sum_by_day$date)
sum_by_day <- cbind(sum_by_day, days)

# display a histogram plot
qplot(days, data=sum_by_day, weight=sumsteps, geom="histogram", binwidth=1)
```

The mean of the total number of steps per day is calculated as:

```{r}
meansteps <- mean(sum_by_day$sumsteps, na.rm=TRUE)
meansteps
```

While the median of the total number of steps per day is calculated as:

```{r}
mediansteps <- median(sum_by_day$sumsteps, na.rm=TRUE)
mediansteps
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?



You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
